Whole-genome pairwise alignment using the hierarchical wavefront algorithm

### 2022-03-30 manuscript/fig_wflamba.tex
figure

\begin{figure}[ht!]
    \begin{subfigure}{\linewidth}
        %\caption{}
        \centering
        % include first image
        \includegraphics[width=1.0\linewidth, trim=-0cm 2cm 0 0cm]{fig/wflambda/chr6_pan_fa_a2fb268_4030258_6a1ecc2_smooth_C4_bad_sorted}
        \label{fig:bad-sorting}
    \end{subfigure}
%	\includegraphics[width=\linewidth]{fig/metrics/chr4.pan.HTTex1.gfa.multiqc_odgi_stats.png}
    \caption{
        PLACEHOLDER IMAGE XXX XXX XXX
    }
    \label{fig:wflambda}
\end{figure}

>>>>

\begin{figure}[ht!]
    \begin{subfigure}{\linewidth}
        %\caption{}
        \centering
        % include first image
        \includegraphics[width=1.0\linewidth, trim=-0cm 2cm 0 0cm]{fig/wflign.png}
        \label{fig:bad-sorting}
    \end{subfigure}
%	\includegraphics[width=\linewidth]{fig/metrics/chr4.pan.HTTex1.gfa.multiqc_odgi_stats.png}
    \caption{
      \textit{wflign} exploration of the alignment of $\sim$560kbp homologus segments of two \textit{Escherichia coli} genomes (GenBank NC\_000913.3 and NC\_008563.1).
      Gray indicates cells explored by WF$\lambda$.
      Blue cells are attempted, failed local WFA alignments, while green cells indicate alignments that match the maximum local alignment score.
      %The traceback (not shown) drives through the 
    }
    \label{fig:wflambda}
\end{figure}

### 2022-03-30 manuscript/main.tex
update simulations
figure

\title[Whole-cromosome pairwise alignment]{Whole-chromosome pairwise alignment using the hierarchical wavefront algorithm}
\author[Guarracino \textit{et~al}.]{
Andrea~Guarracino\,$^{\orcidlink{0000-0001-9744-131X}\text{\sfb 1}}$
Njagi~Mwaniki\,$^{\text{\sfb 2}}$,
Santiago~ Marco-Sola\,$^{\orcidlink{0000-0001-7951-3914}\text{\sfb 3, 4}}$,
Erik~Garrison\,$^{\orcidlink{0000-0003-3821-631X}\text{\sfb 5}*}$
}

@@

\corresp{
$^\ast$To whom correspondence should be addressed.
%$^\dagger$Contributed equally.
}

@@

\abstract{
\textbf{Motivation:}
Pairwise alignment is a critical step in studying whole-genome assemblies and their relationships in an unbiased way.
The time and memory required to compute the pairwise alignments increase quadratically with the sequence length.
This makes it impractical to align very long sequences without applying heuristic approaches to first determine possible synthenic regions in the sequences.
Nevertheless, with the advances in sequencing technologies, new whole-genome assemblies are produced at a high rate, pressing for the development of tools able to align sequences of the order of tens of megabases long. \\
\textbf{Results:}
Here we present wfmash, a new gap-affine pairwise aligner designed to align DNA sequences at a whole-chromosome scale.
wfmash applies a hierarchical implementation of the wavefront alignment algorithm to guide the alignment of very long sequences.
It scales efficiently to large eukaryotic chromosomes, allowing users to perform pairwise alignment of thousands of large genomes using little time and memory. \\
\textbf{Availability:}
wfmash is published as free software under the MIT open source license.
Source code and documentation are available at \url{https://github.com/ekg/wfmash}.
wfmash can be installed via Bioconda \url{https://bioconda.github.io/recipes/wfmash/README.html} or GNU Guix \url{https://github.com/ekg/guix-genomics/blob/master/wfmash.scm}. \\
\textbf{Contact:} \href{egarris5@uthsc.edu}{egarris5@uthsc.edu} \\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.
}

>>>>

\title[Whole-genome pairwise alignment]{Whole-genome pairwise alignment using the hierarchical wavefront algorithm}
\author[Guarracino \textit{et~al}.]{
Andrea~Guarracino\,$^{\orcidlink{0000-0001-9744-131X}\text{\sfb 1}}$,
Njagi~Mwaniki\,$^{\text{\sfb 2}}$,
Santiago~ Marco-Sola\,$^{\orcidlink{0000-0001-7951-3914}\text{\sfb 3, 4}}$,
Erik~Garrison\,$^{\orcidlink{0000-0003-3821-631X}\text{\sfb 5}*}$
}

@@

\corresp{
$^\ast$To whom correspondence should be addressed.
$^\dagger$Contributed equally.
}

@@

\abstract{
\textbf{Motivation:}
Pairwise alignment is a critical step in studying whole-genome assemblies and their relationships in an unbiased way.
The time and memory required to compute the pairwise alignments increase quadratically with sequence length and nucleotide divergence.
This makes it impractical to align very long sequences without applying heuristic approaches to first determine possible synthenic regions in the sequences.
Nevertheless, with the advances in sequencing technologies, new whole-genome assemblies are produced at a high rate, pressing for the development of tools able to align sequences of the order of tens of megabases long. \\
\textbf{Results:}
Here we present wfmash, a new gap-affine pairwise aligner designed to align DNA sequences at a whole-chromosome scale.
wfmash applies a hierarchical implementation of the wavefront alignment algorithm to guide the alignment of very long sequences.
It scales efficiently to large eukaryotic chromosomes, allowing users to perform pairwise alignment of thousands of large genomes using little time and memory. \\
\textbf{Availability:}
wfmash is published as free software under the MIT open source license.
Source code and documentation are available at \url{https://github.com/ekg/wfmash}.
wfmash can be installed via Bioconda \url{https://bioconda.github.io/recipes/wfmash/README.html} or GNU Guix \url{https://github.com/ekg/guix-genomics/blob/master/wfmash.scm}. \\
\textbf{Contact:} \href{egarris5@uthsc.edu}{egarris5@uthsc.edu} \\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.
}
